\Large Generative Dynamics of Supreme Court Citations

### 2020-04-13 Tex/paper.tex
Updated section 4.2 Predictive Performance

\subsubsection{Covariate terms}

@@

\subsubsection{Dependence terms}

@@

Our case for studying legal citations at the directed dyadic level hinges upon the contribution to modeling offered by incorporating network dependence. To quantify this contribution, we use out-of-sample prediction. Predicting out-of-sample offers an unbiased and general purpose way to evaluate the contribution, in terms of model fit, of one or more terms/parameters in a model \citep{jensen2000multiple,ward2010perils}. Unlike in-sample measures of model fit, out-of-sample methods are highly robust in avoiding overfitting, and work when we cannot accurately calculate the value of the likelihood function, as in the current case. Out-of-sample prediction is a common way to evaluate methods for modeling ties in networks, and has recently been applied to TERGMs in particular \citep{desmarais2013forecasting,cranmer2017can}.

In our prediction experiment we randomly split the directed dyads into an 80\% training set and a 20\% test set. The parameters of the model are estimated using the directed dyads in the training set, and the parameters are used to form the conditional probability of a tie for all of the directed dyads in the test set. Directed dyads for which the conditional probability of a tie exceeds 0.5 are predicted to be citations. The experiment is run with the full model, and with a model that excludes all of the dependence terms (i.e., all terms involving reciprocity, in and out stars, and triangles)---the independent dyads model. We run this experiment for 10 iterations. Predictive performance is evaluated with three common and related measures---precision (i.e., the proportion of predicted citations that are actually citations), recall (the proportion of actual citations that are predicted to be citations), and the F1 score (the harmonic mean of precision and recall) \citep[see, e.g.,][for discussion of these measures]{makhoul1999performance}. All three measures are bounded between 0 and 1, with higher scores indicating better performance.

\begin{table}[H]
\centering

\begin{tabular}{rllll}
\hline \hline
& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
 & mean & range & mean & range \\ 
  \hline
precision & 0.6007 & (0.5934, 0.6074) & 0.8666 & (0.8647, 0.8701) \\ 
  recall & 0.1522 & (0.1495, 0.1549) & 0.5969 & (0.5945, 0.6019) \\ 
  F1 score & 0.2429 & (0.2389, 0.2462) & 0.7069 & (0.7048, 0.7107) \\ 
   \hline \hline
\end{tabular}
\caption{The predictive performance of the directed dyadic models with linear trends, over ten 80/20 train/test splits.}
\label{tab:prediction_linear}

We see from Table \ref{tab:prediction_linear} that, based on all three measures, the predictive performance of the model improves dramatically from adding the network dependence terms. The recall of the full model is particularly impressive, indicating that it recovers over half of the actual citations in the test set. This provides clear evidence that the full model, which includes covariates and network dependence terms, represents a more accurate and complete model of the process of citation formation in US Supreme Court opinions.

\begin{table}[H]
\centering
\begin{tabular}{rllll}
\hline \hline
& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
 & mean & range & mean & range \\ 
  \hline
precision & 0.5938 & (0.5844, 0.5996) & 0.8668 & (0.8653, 0.8693) \\ 
  recall & 0.1551 & (0.1515, 0.1575) & 0.5969 & (0.5937, 0.6005) \\ 
  F1 score & 0.246 & (0.2406, 0.249) & 0.707 & (0.7045, 0.7095) \\  
   \hline \hline
\end{tabular}
\caption{The predictive performance of the directed dyadic models with no trends, over ten 80/20 train/test splits.}
\label{tab:prediction_notrend}
\vspace{0.4cm}
\begin{tabular}{rllll}
\hline \hline
& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
 & mean & range & mean & range \\ 
  \hline
precision & 0.6209 & (0.611, 0.6351) & 0.8669 & (0.8612, 0.8711) \\ 
  recall & 0.1511 & (0.1451, 0.1558) & 0.5962 & (0.5884, 0.6011) \\ 
  F1 score & 0.243 & (0.2347, 0.2486) & 0.7065 & (0.7011, 0.7103) \\ 
   \hline \hline
\end{tabular}
\caption{The predictive performance of the directed dyadic models with quadratic trends, over ten 80/20 train/test splits.}
\label{tab:prediction_quadratic}
\end{table}
We also conduct a predictive experiment for the model that does not consider any trends and for the model that does consider quadratic instead of linear trends. The predictive performance results are given in tables \ref{tab:prediction_notrend} and \ref{tab:prediction_quadratic}. By comparing the full model results in tables \ref{tab:prediction_notrend} and \ref{tab:prediction_quadratic} with the results of the full linear trend model in table \ref{tab:prediction_linear} we can conclude all three models perform similarly well.   \\[0,5cm]

@@

\begin{figure}[H]
 \begin{tabular}{cc}

   (a) Reciprocity & (b) Transitivity \\
\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend_linear.pdf} \\

    (c) Activity & (d) Popularity \\
\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend_linear.pdf} \\

\end{tabular}
\caption{Trending effects for dependence effect estimates from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
 \label{fig:coeftrends_endo}
\vspace{-.25cm}
\end{figure}

  \begin{figure}[H]
  \begin{tabular}{cc}
 (e) MQ Score Difference & (f) Ideological Breadth \\
\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mq_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/absdiffmq_coef_trend_linear.pdf} \\

  (g) Same Issue & (h) Majority Size \\
\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/sameissue_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/numberjusticespro_coef_trend_linear.pdf} \\

 (i) Difference in Years & (j) Difference in Years$^2$ \\
\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiff_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiffsquare_coef_trend_linear.pdf} \\

 (k) Overruled Cases & (l) Justice Homophily \\
\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/overruled_coef_trend_linear.pdf} & 
\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/justicehomophily_coef_trend_linear.pdf} \\

 \end{tabular}
\caption{Trending effects for exogenous variables from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
 \label{fig:coeftrends_exo}
\vspace{-.25cm}
\end{figure}

@@

  \begin{figure}[H]
\includegraphics[width=15cm]{justice_popularity}
\caption{Nodeifactor Majority Opinion Writer. Baseline= LDBrandeis }
 \label{number_supporting}
\vspace{-.25cm}
\end{figure}

 \begin{figure}[H]
\includegraphics[width=15cm]{justice_activity}
\caption{Nodeofactor Majority Opinion Writer. Baseline= LDBrandeis }
 \label{number_supporting}
\vspace{-.25cm}
\end{figure}

>>>>

\subsubsection{Covariate terms}\label{covariate_terms}

@@

\subsubsection{Dependence terms}\label{dependence_terms}

@@

Our case for studying legal citations at the directed dyadic level hinges upon the contribution to modeling offered by incorporating network dependence. To quantify this contribution, we use AIC and BIC out-of-sample prediction. Predicting out-of-sample offers an unbiased and general purpose way to evaluate the contribution, in terms of model fit, of one or more terms/parameters in a model \citep{jensen2000multiple,ward2010perils}. Unlike in-sample measures of model fit, out-of-sample methods are highly robust in avoiding overfitting, and work when we cannot accurately calculate the value of the likelihood function, as in the current case. Out-of-sample prediction is a common way to evaluate methods for modeling ties in networks, and has recently been applied to TERGMs in particular \citep{desmarais2013forecasting,cranmer2017can}.

  \begin{figure}[]
\includegraphics[width=15cm]{SCC_AIC_BIC}
\caption{AIC and BIC for the full and the independent model for the time frame 1950-2015 }
 \label{AIC_BIC}
\vspace{-.25cm}
\end{figure}

Figure \ref{AIC_BIC} depicts the AIC and BIC comparison for two different models for each term from 1950 to 2015. The full model is defined as the model that includes all covariate terms and dependence terms described in \ref{covariate_terms} and \ref{dependence_terms}. The independent model only consists of covariate terms, not accounting for any structural dependencies in the networks. We see that the AIC and BIC is considerably smaller for the full model throughout the period of consideration. On average the independent model yields a 5525 point higher AIC and an average 5582 point higher BIC. The maximum difference is reached in 1983 with a difference of 10,990 points for the AIC and 10,930  points for the BIC.

%In our prediction experiment we randomly split the directed dyads into an 80\% training set and a 20\% test set. The parameters of the model are estimated using the directed dyads in the training set, and the parameters are used to form the conditional probability of a tie for all of the directed dyads in the test set. Directed dyads for which the conditional probability of a tie exceeds 0.5 are predicted to be citations. The experiment is run with the full model, and with a model that excludes all of the dependence terms (i.e., all terms involving reciprocity, in and out stars, and triangles)---the independent dyads model. We run this experiment for 10 iterations. Predictive performance is evaluated with three common and related measures---precision (i.e., the proportion of predicted citations that are actually citations), recall (the proportion of actual citations that are predicted to be citations), and the F1 score (the harmonic mean of precision and recall) \citep[see, e.g.,][for discussion of these measures]{makhoul1999performance}. All three measures are bounded between 0 and 1, with higher scores indicating better performance.

%\begin{table}[H]
%\centering

%\begin{tabular}{rllll}
%\hline \hline
%& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
% & mean & range & mean & range \\ 
%  \hline
%precision & 0.6007 & (0.5934, 0.6074) & 0.8666 & (0.8647, 0.8701) \\ 
%  recall & 0.1522 & (0.1495, 0.1549) & 0.5969 & (0.5945, 0.6019) \\ 
%  F1 score & 0.2429 & (0.2389, 0.2462) & 0.7069 & (0.7048, 0.7107) \\ 
%   \hline \hline
%\end{tabular}
%\caption{The predictive performance of the directed dyadic models with linear trends, over ten 80/20 train/test splits.}
%\label{tab:prediction_linear}

%\end{table}

%We see from Table \ref{tab:prediction_linear} that, based on all three measures, the predictive performance of the model improves dramatically from adding the network dependence terms. The recall of the full model is particularly impressive, indicating that it recovers over half of the actual citations in the test set. This provides clear evidence that the full model, which includes covariates and network dependence terms, represents a more accurate and complete model of the process of citation formation in US Supreme Court opinions.

%\begin{table}[H]
%\centering
%\begin{tabular}{rllll}
%\hline \hline
%& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
% & mean & range & mean & range \\ 
%  \hline
%precision & 0.5938 & (0.5844, 0.5996) & 0.8668 & (0.8653, 0.8693) \\ 
%  recall & 0.1551 & (0.1515, 0.1575) & 0.5969 & (0.5937, 0.6005) \\ 
%  F1 score & 0.246 & (0.2406, 0.249) & 0.707 & (0.7045, 0.7095) \\  
%   \hline \hline
%\end{tabular}
%\caption{The predictive performance of the directed dyadic models with no trends, over ten 80/20 train/test splits.}
%\label{tab:prediction_notrend}
%\vspace{0.4cm}
%\begin{tabular}{rllll}
%\hline \hline
%& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
% & mean & range & mean & range \\ 
%  \hline
%precision & 0.6209 & (0.611, 0.6351) & 0.8669 & (0.8612, 0.8711) \\ 
%  recall & 0.1511 & (0.1451, 0.1558) & 0.5962 & (0.5884, 0.6011) \\ 
%  F1 score & 0.243 & (0.2347, 0.2486) & 0.7065 & (0.7011, 0.7103) \\ 
%   \hline \hline
%\end{tabular}
%\caption{The predictive performance of the directed dyadic models with quadratic trends, over ten 80/20 train/test splits.}
%\label{tab:prediction_quadratic}
%\end{table}
%We also conduct a predictive experiment for the model that does not consider any trends and for the model that does consider quadratic instead of linear trends. The predictive performance results are given in tables \ref{tab:prediction_notrend} and \ref{tab:prediction_quadratic}. By comparing the full model results in tables \ref{tab:prediction_notrend} and \ref{tab:prediction_quadratic} with the results of the full linear trend model in table \ref{tab:prediction_linear} we can conclude all three models perform similarly well.   \\[0,5cm]

@@

%\begin{figure}[H]
% \begin{tabular}{cc}

%   (a) Reciprocity & (b) Transitivity \\
%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mutual_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/triangle_coef_trend_linear.pdf} \\

%    (c) Activity & (d) Popularity \\
%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/o2star_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/i2star_coef_trend_linear.pdf} \\

%\end{tabular}
%\caption{Trending effects for dependence effect estimates from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
% \label{fig:coeftrends_endo}
%\vspace{-.25cm}
%\end{figure}

%  \begin{figure}[H]
%  \begin{tabular}{cc}
% (e) MQ Score Difference & (f) Ideological Breadth \\
%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/mq_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/absdiffmq_coef_trend_linear.pdf} \\

%  (g) Same Issue & (h) Majority Size \\
%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/sameissue_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/numberjusticespro_coef_trend_linear.pdf} \\

% (i) Difference in Years & (j) Difference in Years$^2$ \\
%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiff_coef_trend_linear.pdf} & \includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/yeardiffsquare_coef_trend_linear.pdf} \\

% (k) Overruled Cases & (l) Justice Homophily \\
%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/overruled_coef_trend_linear.pdf} & 
%\includegraphics[width = 0.475\textwidth, trim= 0.1cm 1cm 0.5cm .45cm,clip=true]{images/justicehomophily_coef_trend_linear.pdf} \\

% \end{tabular}
%\caption{Trending effects for exogenous variables from the c-TERGM. The solid lines plot the point estimates, and the shaded areas span 95\% confidence intervals. }
% \label{fig:coeftrends_exo}
%\vspace{-.25cm}
%\end{figure}

@@

  \begin{figure}[H]
\includegraphics[width=16cm]{SCC_Results_1}
\caption{ERGM results for the dependence terms. Circles indicate a p-value smaller than 0.05, squares a p-value between 0.05 and 0.1 and triangles a p-value greater than 0.1. }
 \label{SCC_results_1}
\vspace{-.25cm}
\end{figure}

 \begin{figure}[H]
\includegraphics[width=16cm]{SCC_Results_2}
\caption{ERGM results for the covariate terms. Circles indicate a p-value smaller than 0.05, squares a p-value between 0.05 and 0.1 and triangles a p-value greater than 0.1.}
 \label{SCC_results_2}
\vspace{-.25cm}
\end{figure}
